---
title: "Emulating LMDZ SCM output"
output: html_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(fig.path='figure/lmdz/', 
                      fig.height=6, fig.width=10,
                      cache.path='_knitr_cache/emulate-lmdz/')
set.seed(14112017)
```

In this example we use primarily R and Stan files from **BuildEmulator** directory to construct and validate a GP emulator. 

First we start with loading the necessary packages, functions and compilation of Stan files that are necessary to construct a GP emulator.
```{r results='hide', message=FALSE, warning=FALSE}
setwd('..')
source('BuildEmulator/BuildEmulator.R')
```

We proceed to loading the data file `lmdzAWave1.RData`. The emulator inputs are saved in columns 1 to 6 (on $[-1, 1]$ scale) of the object `tData` together with the simulator response saved as a last column of a data frame.

```{r}
load('../Demonstrations/lmdzAWave1.RData')
head(tData)
```

We specify a number and a vector of metrics of interest (computer model output), a number and vector of candidates (potential emulator inputs). Notice that we have an extra `Noise` variable, which is considered as a potential predictor and plays an important part in deriving the form of a regression function $h(\boldsymbol{x})$.

```{r}
metric = c('SCMdata')
nmetric = length(metric)
cands = names(tData)[-which(names(tData) == metric)]
param = cands[-which(cands == "Noise")]
nparam = length(param)
```

We visualize the model response behaviour against model inputs standardized on $[-1, 1]$ scale via scatter plots:

```{r plot-lmdz-scatter}
if(nparam*nmetric<2){ par(mfrow = c(1, 1), mar=c(4, 4, 1, 1))
} else if(nparam*nmetric <= 4){ par(mfrow = c(2, 2), mar=c(4, 4, 1, 1))
} else if(nparam*nmetric <= 9){ par(mfrow = c(3, 3), mar=c(4, 4, 1, 1))
} else if(nparam*nmetric <= 16){ par(mfrow = c(4, 4), mar=c(4, 4, 1, 1))
} else if(nparam*nmetric <= 25){ par(mfrow = c(5,5), mar=c(4, 4, 1, 1)) }
for ( iparam in 1:nparam ) {
  for (j in 1:nmetric) {
    plot(tData[,iparam],tData[,nparam+j+1],col=2,xlab=cands[iparam],ylab=metric[j],
         ylim=range(tData[,nparam+j+1]))
  }
}
```

## Build the emulator

We build the emulator for the output variable (metric) `SCMdata`. In this particular example we have just one metric of interest `nmetric=1`, however it is possible to extend `InitialBasisEmulators` to $q>1$ uncorrelated metrics. `InitialBasisEmulators` constructs a GP emulator for one at a time metric of interest in two major steps:

1. Derive the form of a regression function $h(\boldsymbol{x})$ by using a stepwise regression procedure. We use function `EMULATE.lm`.
2. Construct a full GP emulator with a function `EMULATE.gpstan`.

```{r results='hide', message=FALSE, warning=FALSE}
listnew <-lapply(1:nmetric,function(k) names(tData[1:nparam]))
myem.lm <-InitialBasisEmulators(tData=tData,HowManyEmulators=nmetric,additionalVariables = listnew)
```

The output of a function is a list where each member of a list corresponds to an object `StanEmulator` obtained for each metric of interest.

```{r}
names(myem.lm[[1]])
```

For example, we could consider a regression function $h(\boldsymbol{x})$ used in the definition of the mean function for a GP emulator

```{r}
summary(myem.lm[[1]]$linModel)
```

## Diagnostics for a GP emulator

Since we are operating with Stan, it is crucial to check the convergence of posterior samples obtained for GP emulator hyperparameters.
```{r plot-lmdz-traceplot}
traceplot(myem.lm[[1]]$StanModel)
```

We also validate the performance of GP emulator by considering the Leave One Out (LOO) diagnostics plots. The black dots and error bars shows the predictions $\pm2\times$ standard deviation prediction intervals from the leave one out emulators, whilst the green/red dots are the true model output coloured by whether or not the true lies within the error bar.
```{r plot-lmdz-LOO}
for (i in 1:nmetric) {
  tLOOs1 <- LOO.plot(StanEmulator = myem.lm[[i]] , ParamNames=names(tData)[1:nparam])
}
```


